---
title: "Arrow Experiments"
output: html_notebook
---

This notebook is for experimenting with the `arrow` R package.

```{r setup}
# Packages ----
library(arrow)
library(lobstr)
library(vroom)
library(tidyverse)

# Utils ----
write_chunk_data <- function(data_path, output_dir, chunk_size = 10000000) {
  if (!fs::dir_exists(output_dir)) fs::dir_create(output_dir)
  data_name <- fs::path_ext_remove(fs::path_file(data_path))
  chunk_num <- 0
  data_chunk <- vroom::vroom(data_path, n_max = chunk_size)
  data_names <- names(data_chunk)
  arrow::write_parquet(data_chunk, fs::path(output_dir, glue::glue("{data_name}-{chunk_num}.parquet")))
  while (nrow(data_chunk) == chunk_size) {
    chunk_num <- chunk_num + 1
    data_chunk <- vroom::vroom(data_path, skip = chunk_num * chunk_size, n_max = chunk_size, col_names = data_names)
    arrow::write_parquet(data_chunk, fs::path(output_dir, glue::glue("{data_name}-{chunk_num}.parquet")))
  }
}
```

## Data
Data was downloaded from [NYC OpenData](https://data.cityofnewyork.us/browse?q=taxi) and contains details about NYC taxi trips. The following code takes raw .csv files and chunks them into smaller parquet files so the entire dataset can be queried from within R using `arrow`.
```{r create-data}
if (length(list.files("data/parquet")) == 0) {
  csvs <- fs::dir_ls("data", glob = "*.csv")
  if (length(csvs) == 0) stop("No csv files found in data/")
  walk(csvs, write_chunk_data, "data/parquet")
}  
fs::dir_ls("data/parquet")
```

## Arrow
Use arrow to read in data from directory of parquet files.
```{r read-data}
taxi_a <- open_dataset("data/parquet")
class(taxi_a)
obj_size(taxi_a)
```

Compare object size to a single file read into R from parquet
```{r single-file-size}
taxi_0 <- read_parquet("data/parquet/taxi-0.parquet")
obj_size(taxi_0)
rm(taxi_0)
```

Column names
```{r col-names}
names(taxi_a)
```

## Arrow + dplyr

Number of observations
```{r, eval = FALSE}
taxi_a %>% 
  tally()
```
**Consumes an incredible amount of RAM**

```{r}
taxi_a %>% 
  count(VendorID) %>% 
  summarize(total_rows = sum(n))
```

Count of passengers
```{r}
taxi_a %>% 
  count(passenger_count)
```

Count of VendorIDs
```{r}
taxi_a %>% 
  count(VendorID)
```

Average route distance and amount by passenger count
```{r summarize}
taxi_a %>% 
  group_by(passenger_count) %>% 
  summarize(n_obs = n(),
            avg_distance = mean(trip_distance, na.rm = TRUE),
            avg_charge = mean(total_amount, na.rm =TRUE))
```

Creating new columns with mutate
```{r mutate}
taxi_a %>% 
  mutate(p1 = passenger_count + 1) %>% 
  count(p1)
```
**Doesn't run**

Filtering rows
```{r filter}
taxi_f <- taxi_a %>% 
  filter(passenger_count == 1)

class(taxi_f)
```

```{r}
taxi_f %>% 
  summarize(n_obs = n(),
            avg_distance = mean(trip_distance, na.rm =TRUE),
            avg_charge = mean(total_amount, na.rm = TRUE))
```

```{r}
taxi_a %>% 
  filter(passenger_count > 5) %>% 
  summarize(n_obs = n(),
            avg_distance = mean(trip_distance, na.rm =TRUE),
            avg_charge = mean(total_amount, na.rm = TRUE))
```
Collect results from a query directly back into R
```{r}
taxi_a %>% 
  filter(passenger_count >= 7) %>% 
  collect()
```

